---
title: "Time Series Analysis"
author: "IO"
date: "09 02 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(xts)
library(dplyr)
library(weatherData)
library(ggplot2)
```

## Dataset

```{r}
print(Nile)
nile <- as.xts(Nile)
head(nile)
```

```{r}
plot(Nile, 
     xlab = "Year",
     ylab = "River Volume (1e9 m^{3})",
     main = "Annual River Nile Volume at Aswan, 1871-1970",
     type = "s")
```

```{r}
is.ts(nile)
is.xts(nile)
start(nile)     #starting date of the observations
frequency(nile) #how many observations each day
end(nile)       #last date of observations
deltat(nile)    #time interval between observations in days (how many observations in a day)
```

## London weather data form 2013

We should tranform the data into an xts object

```{r}
is.ts(London2013)
london <- as.xts(x = London2013$Temperature,
       order.by = lubridate::as_datetime(London2013$Time))

periodicity(london)
```

```{r}
plot(london)
```

```{r}
plot(london["201301/201303"])
plot(xts::first(london, "3 months"))
```

Differences between every 10 observation

```{r}
head(
  merge(london,
        diff(london, 
             lag = 10)),
  20
)
```

Mean Temp of every month

```{r}
plot(apply.monthly(london,
              FUN = "mean"),
     type = "s",
     main = "month")
```

OHLC Values

```{r}
to.period(london,
          period = "months")
```

## ARIMA models

ARIMA: AutoRegressive, Integrated Moving Average

We can use ARIMA models to specify white noise (WN) models.

```{r}
# Stimulate a WN model with 50 obervations
WN <- arima.sim(model = list(order = c(0,0,0)),
                n = 50)
# This series has a default value of mean and sd, with are 0 and 1
head(WN)

ts.plot(WN)
```

```{r}
# Set the mean and sd of a stimulated WN model
WN_1 <- arima.sim(model = list(order =  c(0,   #Autoregressive order
                                          0,   #Order of integration (differencing)
                                          0)), #Moving average order
                  n = 50,
                  mean = 4,
                  sd = 2)

head(WN_1)

# We can see this plot has a bigger mean and more variation
ts.plot(WN_1)
```

### Estimating the white noise in a time series

```{r}
arima(WN_1, 
      order = c(0,0,0))
```

The value that is given as the blank row in under intercept is mean.  
s.e. is the strandard error.  
sigma^2 is indicating the variance in the model. The estimated standard deviation is the square root of this value.

```{r}
mean(WN_1)
var(WN_1)
```

